Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-08-27 grove-paper.tex
Add labels to all sections.  Update the multi-user sections with Alice and Bob and references to the new figures.

\section{Introduction}
\label{sec:introduction}

@@

\subsection{Contributions and Paper Organization}
\label{sec:paper-organization}

\section{Grove By Example}
\label{sec:grove-by-example}

TODO: "run through several senarios though in practice these occur less frequently than portrayed here"

TODO: Alice and Bob and narrative continuity

@@

This section introduces collaborative structure editing in Grove by example. 
We begin in \autoref{sub:Program Representation} by introducing basic edit actions
and the underlying graph representation of expressions with holes.
We will then discuss scenarios that arise in the collaborative setting in 
\autoref{sub:Collaborative Structure Editing}.

TODO: gray-scale unchanged parts of diagrams; and dashes for removed parts?

TODO: typewritter font in figure graphs

We present this at two levels.
One is the underlying graph model, which
involves explicit identifiers, graph edges, indexes, and add and delete
operations.
The other is the interface a user might see,
which omits these details unless needed.
Throughout these examples the graph model is presented
as a graph with.

@@

The state for each editor instance in Grove consists of an expression containing holes along with a cursor.
For example, \autoref{fig:A simple tree with a hole} shows the arithmetic expression~\texttt{x * \hole}, which
has a hole on the right side of the times operator.

Internally, Grove represents an expression like this as a directed graph with a distinguished root vertex.
For example, \autoref{fig:A simple tree with a hole-b} shows the graph corresponding to the term in \autoref{fig:A simple tree with a hole-a}.
Each vertex represents a term and is labeled with a globally unique identifier and a term constructor.
In \autoref{fig:A simple tree with a hole}, \vertex{2} and \vertex{4} have the constructors \texttt{times} and \texttt{var(x)}, respectively.

Each term constructor has an associated set of child positions.
The~\texttt{*} constructor has positions for~\texttt{left} and~\texttt{right} children
and the~\texttt{var} constructors have no child positions but takes an argument naming the particular
variable referenced~(e.g.,~\texttt{x} in \texttt{var(x)}).
The root vertex has a distinguished identifier~0 and the constructor \textbullet with the single child position \texttt{root}.

Each edge is labeled with a globally unique identifier (e.g.,~1 and~3 in the figure) and a child position (e.g., \texttt{left} and \texttt{root} in the figure).
An edge indicates that the destination vertex is a child of the origin vertex at the given position.
%Visually, we indicate the position of an edge by the location of its origin.
For the sake of presentation clarity, we reserve odd identifiers for vertexes and even identifiers for edges.

@@

Note that, for the purposes of this paper, we represent identifiers and number literals
indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.

@@

\subsubsection{Additions}

@@

We model edits to the code as sequences of additions and deletions of edges to this graph.
For example, suppose the user adds~\texttt{y} as the right child of \vertex{2} in
\autoref{fig:A simple tree with a hole}
so that the expression becomes \texttt{x * y}.
This corresponds to creating a vertex for the variable reference~\texttt{y} and adding an edge
to that vertex from \vertex{2}.
\autoref{fig:Addition} shows the result, where the new vertex has identifier~6.

TODO: (though as we show in \autoref{sec:formalism} creating a vertex is implicit in edge creation)

@@

\subsubsection{Changes}

@@

Next suppose we want to replace~\texttt{y} with~\texttt{z} so that the code becomes~\texttt{x * z}.
This is accomplished by first deleting \edge{5}.
This results in \autoref{fig:Change-a}, which represents the expression~\texttt{x * \hole{}}.
After that deletion, we add a new vertex to represent~\texttt{z} and an
edge from \vertex{2} to that vertex.
This results in the bottom graph in \autoref{fig:Change-b},
which represents the expression~\texttt{x * z}.

Note that this edge has a new identifier distinct from the one before.
An edge once deleted cannot be restored.  Instead a new edge (with a new id) is created.

@@

Explain orphans

\subsubsection{Wrapping Code}

@@

TODO: mention that naive add that creates a multi-child is discussed later

Next suppose a user inserts cod at a position already occupied by some code.
For example, suppose, the user types~\texttt{+} when the cursor is on \vertex{2}.
If this were a hole, this would simply put a vertex with the~\texttt{+} constructor
as a child of \vertex{0}.

If the user requests that a plus be inserted,
the editor may notice that there is something already in that hole.
A naive editor could certainly add the plus by creating a new
edge from the body index of the lambda to a freshly created \texttt{plus} vertex.
This would result in a conflict between the \texttt{plus} and the existing \texttt{var}
as shown in Figure~REF:TODO.
A better user experience
would be interpret the request for a plus to mean to
take whatever (if anything) is in the cursor
and place it as a child of the plus.
This \emph{wraps} the existing \texttt{var} in the \texttt{plus}
and results in the code~\texttt{$\lambda$ x . x + \hole}.
(This is similar to the behavior in other structure editors such as in CITE:TODO.)
Translating this into edge deletes and adds results in the sequence of graphs in \autoref{fig:Wrapping}.
Note that this is a layer above

Traditional diff: no relation (and indent might change)

\subsubsection{Move}

@@

For example, consider \autoref{fig:Move}, which represents
the movement of~\texttt{1 * 2} from the left child of \texttt{plus}
to its right child.  The top graph is the code before the move.
First, \edge{7} is deleted, which results in the middle graph
where the rooted code is now~\texttt{\hole{} + \hole},
and \vertex{6} is orphaned.

Next \edge{16} is added, which results in the bottom graph
where the rooted code is now~\texttt{\hole{} + 1 * 2},
and \vertex{6} is no longer orphaned.

@@

TODO: mention that id numbers are global

\subsubsection{Editing Different Parts of the Code}

@@

When editing different parts of the code

For example, suppose there are users, A and B,
that are synchronized with each other and
have the code~\texttt{3 + 4}.
That is to say, they both have
the top graphs in \autoref{fig:Editing Different Parts of the Code}.

Next, suppose user A changes the~\texttt{3} to a~\texttt{13},
and user B changes the~\texttt{4} to a~\texttt{14}.
Before transmitting their edits to each other,
user A thus has left graph and user B has the right graph.
For user A, this involves deleting \edge{7} and creating \edge{14}.
For user B, this involves deleting \edge{10} and creating \edge{18}.

To merge these two graphs each user simply replays the edge deletions and additions
that the other user performed.

Both user then have the bottom graph and are again synchronized with each other.

Note that aside from transmitting the edge deletions and additions,
no coordination between A and B is needed to coordinate the merge.
For example, suppose that A and B did not transmit their edits to each other
but to a third user, C, that has the original graph at the top of \autoref{fig:Editing Different Parts of the Code}.
User C can replay the edits from A and B in any order and will always get
the graph at the bottom of \autoref{fig:Editing Different Parts of the Code}.

Stated more formally, our edit model obeys commutativity
as shown in \autoref{fig:Commutativity}.
For any two sequences of edits $\vec{\alpha}$ and $\vec{\alpha}^\prime$,
applying them in either order produces the same results.

@@

\subsubsection{Editing Nested Parts of the Code}

(Similar to how git merged across moves.
A notable difference though is that \texttt{git} detects
file moves by structural similarity rather than
our nominal equality.
(We exploit this in Subsection~REF:TODO).

@@

Consider \autoref{fig:Editing Nested Parts of the Code}.
Alice and Bob both start with the top graph, which represents the code~\texttt{2 * x + \hole{}}.
Alice changes the~\texttt{2} to a~\texttt{3}, which results in the left graph.
Bob moves the~\texttt{2 * x} from the left of the plus to the right of the plus
by deleting \edge{7} and adding \edge{16}, which results in the right graph.
Since vertex and edge identifiers are preserved by these,
both Alice's and Bob's edits can be applied to the other's graph,
which results in both Alice and Bob having the bottom graph.

As with Subsection~SEC:TODO, this merge does not require coordination
between Alice and Bob.  They could independently send their edits
to a third party, Charley, who could replay their edits in any order
and will always get the bottom graph.

A consequence of this design is that we allow users to edit
parts of the code that have been deleted.  For example,
Alice might receive Bob's deletion of \edge{7} before making her
edits and before receiving Bob's adding of \edge{16}.
Her edits would then be within the code deleted by Bob.

@@

\subsubsection{Multi-child conflicts}

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{*}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
    "$\texttt{var(w)}_{16}$" [> "right", > "17"']
   },
   {
    "$\texttt{var(x)}_{20}$" [> "right", > "21"']
   }
  }
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Editing Nested Parts of the Code-c} after Alice ... TODO}
\label{fig:Multi-child-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{*}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
    "$\texttt{var(w)}_{16}$" [> "right", > "17"']
   },
   {
    "$\texttt{app}_{22}$" [> "right", > "23"'] -> {
     "$\texttt{var(f)}_{24}$" [> "fun", > "25"'],
     "$\texttt{var(y)}_{26}$" [> "arg", > "27"']
    }
   }
  }
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Editing Nested Parts of the Code-c} after Bob ... TODO}
\label{fig:Multi-child-b}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{*}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
    "$\texttt{var(w)}_{16}$" [> "right", > "17"']
   },
   {
    "$\texttt{var(x)}_{20}$" [> "right", > "21"'],
    "$\texttt{app}_{22}$" [> "right", > "23"'] -> {
     "$\texttt{var(f)}_{24}$" [> "fun", > "25"'],
     "$\texttt{var(y)}_{26}$" [> "arg", > "27"']
    }
   }
  }
 }
};
\end{tikzpicture}
\caption{The result of merging \autoref{fig:Multi-child-a} with \autoref{fig:Multi-child-b}}
\label{fig:Multi-child-c}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Multi-child conflicts}
\label{fig:Multi-child}
\end{figure}

@@

Finally, we address genuine conflicts between edits.
Consider the top graph of \autoref{fig:Multi-child}.
Suppose Alice inserts the number literal~\texttt{2} into the body of the lambda
while Bob inserts the function application~\texttt{f x} also into the body of the lambda.
These results in the left and right graphs of \autoref{fig:Multi-child} respectively.
In particular pay attention to the edges going out of the \texttt{body} index of \vertex{3}.
In the left graph it is \edge{10}, which points to a number literal,
and in the right graph it is \edge{13}, which points to a function application.

We merge these by including \emph{both} edges as edges going out of the \texttt{body} index
of \vertex{3}.
This results in the bottom graph of \autoref{fig:Multi-child}.
The conflict between Alice's and Bob's changes is
represented by the fact that there are multiple edges going out of one index (i.e. the \texttt{body}
index of \vertex{3}).
In the graph representation, we flag this conflict by highlighting its edges
in red.
In the textual representation, we flag this conflict with the notation~\texttt{\{ 2 | f x \}}.
This is a conflict that much eventually be resolved by the user by deleting
one of the edges and editing the other to be the merged value.
For example, one could delete \edge{10} and then edit the~\texttt{x}
to be~\texttt{x + 2}.
Note that users can continue to edit the code while in this conflicted state.
It just must be resolved before the the code can be run.\footnote{
It may be possible to develop evaluation models that allow these sorts of conflicts,
but that is beyond the scope of this paper.}

@@

\subsubsection{Multi-parent conflicts}

@@

@@

\subsubsection{Cycles}

@@

@@

\section{Formalism}
\label{sec:formalism}

@@

\subsection{Syntax}

@@

\subsection{Convergent Graphs}

\subsubsection{Graphs}

@@

\subsubsection{Graph Actions}
A graph action is a +E or -E.

@@

\subsection{Agda Mechanization}
\label{sub:agda-mechanization}

@@

\subsection{Type System}
\label{sub:type-system}

@@

\subsection{Edit Actions}
Mapping to graph actions

@@

\section{Implementation}
\label{sec:implementation}
GRV -- how it is implemented, how it connects to the formalism, describe the graph,

@@

\section{Related Work}
\label{sec:related-work}

@@

\section{Discussion and Conclusion}
\label{sec:conclusion}

@@

\subsection{Variable names, strings, and numbers}

@@

%% Appendix
\appendix
\section{Appendix}

>>>>

\section{Introduction}
\label{sec:Introduction}

@@

\subsection{Contributions and Paper Organization}
\label{sec:Contributions and Paper Organization}

\section{Grove By Example}
\label{sec:Grove By Example}

This section introduces collaborative structure editing in Grove.
We begin in \autoref{sub:Program Representation} by covering how we represent expressions with holes as graphs,
\autoref{sub:Single User Edits} explores edits by a single user, Alice,
and \autoref{sub:Collaborative Structure Editing} explores edits by multiple users, Alice and Bob,
each of whom is editing their own instance of a program.

@@

TODO: gray-scale unchanged parts of diagrams; and dashes for removed parts? describe semantic vs presentational in diagrams

@@

%We represent expression contain holes as part of the editor state 
The \textit{program state} for each editor instance in Grove consists of a graph that represents the program being edited.
This can be presented to the user as an expression containing holes.
For example, \autoref{fig:A simple tree with a hole-b} shows one such graph,
and \autoref{fig:A simple tree with a hole-a} shows its corresponding expression,~\texttt{x * \hole}, which
has a hole on the right side of the times operator.

%We present this from two perspectives.
%One is the underlying graph model, the other is the interface a user might see, which omits these details unless needed.

%which involves explicit graph edges, positions, and add and delete operations.

Program state is a directed graph with a distinguished root vertex.
%For example, \autoref{fig:A simple tree with a hole-b} shows the graph corresponding to the term in \autoref{fig:A simple tree with a hole-a}.
Each vertex represents a term and is labeled with a globally unique identifier and a term constructor.
In \autoref{fig:A simple tree with a hole-b}, \vertex{2} and \vertex{4} have the constructors~\texttt{*} and \texttt{var(x)}, respectively.
The root vertex has the distinguished identifier~0 and the constructor~\textbullet.

Each term constructor has an associated set of child positions.
The~\texttt{*} constructor has positions for~\texttt{left} and~\texttt{right} children
and the~\texttt{var} constructors have no child positions but takes an argument naming the particular
variable referenced~(e.g.,~\texttt{x} in \texttt{var(x)}).\footnote{Note that,
for the purposes of this paper, we represent identifiers and number literals
indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.}
The root vertex constructor,~\textbullet, has the single child position \texttt{root}.

Each edge is labeled with a globally unique identifier (e.g.,~1 and~3 in \autoref{fig:A simple tree with a hole-b}) and
a child position (e.g., \texttt{left} and \texttt{root} in \autoref{fig:A simple tree with a hole-b}).
An edge indicates that the destination vertex is a child of the origin vertex at the given position.
%Visually, we indicate the position of an edge by the location of its origin.
For the sake of presentation clarity, we reserve odd identifiers for vertexes and even identifiers for edges.

@@

\marginpar{START REV HERE}

@@

\subsubsection{Additions}
\label{sub:Additions}

@@

TODO: Actions act at a cursor.  Each instance also has a cursor that it controls, but this is not part of the graph representation.

Alice starts with the code in \autoref{fig:A simple tree with a hole},
She moves her cursor to the hole in~\texttt{x * \hole} and enters the variable~\texttt{y}.
We model edits as sequences of additions and deletions of edges to the graph.
Alice's edit corresponds to creating a vertex for the variable reference~\texttt{y} and adding an edge
to that vertex from \vertex{2}.
\autoref{fig:Addition-b} shows the result, where the new vertex has identifier~6.
This graph represents the expression \texttt{x * y}.

TODO: (though as we show in \autoref{sec:Formalism} creating a vertex is implicit in edge creation)

@@

\subsubsection{Changes}
\label{sub:Changes}

@@

Next Alice wants to replace~\texttt{y} with~\texttt{z} so that the code becomes~\texttt{x * z}.
This is modeled by first deleting \edge{5}.
This results in \autoref{fig:Change-a}, which represents the expression~\texttt{x * \hole{}}.
After that deletion, we add a new vertex to represent~\texttt{z} and an
edge from \vertex{2} to that vertex.
This results in the bottom graph in \autoref{fig:Change-b},
which represents the expression~\texttt{x * z}.

Note that this edge has a new identifier distinct from the one before.
In our model, once an edge with a particular identifier is deleted, it cannot be recreated.
Instead a new edge with a new identifiers is created.

@@

TODO: Explain orphans

\subsubsection{Wrapping Code}
\label{sub:Wrapping Code}

@@

Next, Alice types~\texttt{+} when the cursor is on \vertex{2}.
If this were a hole, this would simply put a vertex with the~\texttt{+} constructor
as a child of \vertex{0}.
However, this position is already occupied by some code.

A naive editor could add the~\texttt{+} by creating a new
edge from the body index of the lambda to a freshly created \texttt{plus} vertex.
This would result in a conflict between the \texttt{+} and the existing \texttt{var}
as shown in Figure~REF:TODO.

How to handle multiple children in the same position is explored in Sec~REF:TODO,
but for this sort of edit, a better user experience
would be interpret the request for a plus to mean to
take whatever (if anything) is in the cursor
and place it as a child of the plus.
This \emph{wraps} the existing \texttt{var} in the \texttt{plus}
and results in the code~\texttt{(x * z) + \hole}.
(This is similar to the behavior in other structure editors such as in CITE:TODO.)
Translating this into edge deletes and adds results in the sequence of graphs in \autoref{fig:Wrapping}.

TODO: "Note that the logic of to insert or wrap is a layer above graph actions."

TODO: Traditional diff: no relation (and indent might change)

\subsubsection{Move}
\label{sub:Move}

@@

For example, consider \autoref{fig:Move}, which represents
the movement of~\texttt{1 * 2} from the left child of \texttt{plus}
to its right child.
\autoref{fig:Editing Nested Parts of the Code-c} is the code before the move.
First, \edge{7} is deleted, which results in \autoref{fig:Move-a}
where the rooted code is now~\texttt{\hole{} + \hole{}},
and \vertex{2} is orphaned.
Next \edge{11} is added, which results in the bottom graph
where the rooted code is now~\texttt{\hole{} + 1 * 2},
and \vertex{2} is no longer orphaned.

@@

TODO: mention that id numbers are global (or maybe use A1 and B1 for different users?)

We assume multiple instances

\subsubsection{Editing Different Parts of the Code}
\label{sub:Editing Different Parts of the Code}

@@

To start, Alice and Bob are synchronized with each other and
both have the graph in \autoref{fig:Move-b}, which
represents the code~\texttt{\hole{} + (x + z)}.

Alice changes the~\texttt{x} to a~\texttt{u},
and Bob changes the~\texttt{z} to a~\texttt{v}.
Before transmitting their edits to each other,
Alice thus has the graph in \autoref{fig:Editing Different Parts of the Code-a}
and Bob has the graph in \autoref{fig:Editing Different Parts of the Code-b}.
For Alice, this involves deleting \edge{3} and creating \edge{13}.
For Bob, this involves deleting \edge{5} and creating \edge{15}.

Alice and Bob then transmit their graph actions (i.e., edge additions and deletions) to each other
and applies the other's graph actions to their own copy of the graph.
By design, graph actions commute, so this results in both Alice and Bob having the graph in
\autoref{fig:Editing Different Parts of the Code-c}.

Note that aside from transmitting the edge deletions and additions,
no coordination between Alice and Bob is needed to coordinate the merge.
For example, suppose that Alice and Bob did not transmit their edits to each other
but to a third user, Chris, that has the original graph at the top of \autoref{fig:Move-b}.
Chris can apply the edits from Alice and Bob in any order and will always get
the graph in \autoref{fig:Editing Different Parts of the Code-c}.

Stated more formally, our edit model obeys the commutativity shown in \autoref{fig:Commutativity}.
For any two sequences of edits $\vec{\alpha}$ and $\vec{\alpha}^\prime$,
applying them in either order produces the same results.

@@

\subsubsection{Editing Nested Parts of the Code}
\label{sub:Editing Nested Parts of the Code}

@@

After the edits in the previous subsection synchronize,
Alice changes the~\texttt{v} to a~\texttt{w}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-a}.
On the other hand, Bob moves the~\texttt{u * v} from the right of the plus to the left of the plus
by deleting \edge{11} and adding \edge{19}, which results in the graph in \autoref{fig:Editing Nested Parts of the Code-b}.
Since vertex and edge identifiers are preserved by these,
both Alice's and Bob's edits can be applied to the other's graph,
which results in both Alice and Bob having the graph in \autoref{fig:Editing Nested Parts of the Code-c}.

As with \autoref{sub:Editing Different Parts of the Code}, this merge does not require coordination
between Alice and Bob.  They could independently send their edits
to a third party, Charley, who could replay their edits in any order
and will always get the graph in \autoref{fig:Editing Nested Parts of the Code-c}.

A consequence of this design is that we allow users to edit
parts of the code that have been deleted.  For example,
Alice might receive Bob's deletion of \edge{11} before making her
edits and before receiving Bob's addition of \edge{19}.
Her edits would then be within the code deleted by Bob.

@@

TODO: put this somewhere.
(Similar to how git merged across moves.
A notable difference though is that \texttt{git} detects
file moves by structural similarity rather than
our nominal equality.
(We exploit this in Subsection~REF:TODO).

\subsubsection{Multi-child conflicts}
\label{sub:Multi-child conflicts}

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{*}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
    "$\texttt{var(w)}_{16}$" [> "right", > "17"']
   },
   {
    "$\texttt{var(x)}_{20}$" [> "right", > "21"']
   }
  }
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Editing Nested Parts of the Code-c} after Alice ... TODO}
\label{fig:Multi-child-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{*}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
    "$\texttt{var(w)}_{16}$" [> "right", > "17"']
   },
   "$\texttt{var(y)}_{22}$" [> "right", > "23"']
  }
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Editing Nested Parts of the Code-c} after Bob ... TODO}
\label{fig:Multi-child-b}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{*}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
    "$\texttt{var(w)}_{16}$" [> "right", > "17"']
   },
   {
    "$\texttt{var(x)}_{20}$" [> "right", > "21"'],
    "$\texttt{var(y)}_{22}$" [> "right", > "23"']
   }
  }
 }
};
\end{tikzpicture}
\caption{The result of merging \autoref{fig:Multi-child-a} with \autoref{fig:Multi-child-b}}
\label{fig:Multi-child-c}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Multi-child conflicts}
\label{fig:Multi-child}
\end{figure}

@@

Finally, we address our first kind of genuine conflict between edits.
Alice inserts the variable~\texttt{x} to the right side of the~\texttt{+}
while Bob add the variable~\texttt{y} also to the right side of the~\texttt{+}
These results in the graphs in \autoref{fig:Multi-child-a} and \autoref{fig:Multi-child-b}, respectively.
In particular pay attention to the edges going out of the \texttt{right} index of \vertex{8}.
In \autoref{fig:Multi-child-a} it is \edge{21}, which points to the variable~\texttt{x},
and in \autoref{fig:Multi-child-b} it is \edge{23}, which points to the variable~\texttt{y}.

We merge these by including \emph{both} edges as edges going out of the \texttt{right} index
of \vertex{8}.
This results in \autoref{fig:Multi-child-c}.
The conflict between Alice's and Bob's changes is
represented by the fact that there are multiple edges going out of one index (i.e. the \texttt{right}
index of \vertex{8}).
In the graph representation, we flag this conflict by highlighting its edges
in red (TODO).
In the textual representation, we flag this conflict with the notation~\texttt{\{ x | y \}}.
This is a conflict that must eventually be resolved by the user by deleting
one of the edges and editing the other to be the merged value.
For example, one could delete \edge{23} and then edit the~\texttt{x}
to be~\texttt{x * y}.
Note that users can continue to edit the code while in this conflicted state.
It just must be resolved before the the code is run.\footnote{It
may be possible to develop evaluation models that allow these sorts of conflicts,
but that is beyond the scope of this paper.}

@@

\subsubsection{Multi-parent conflicts}
\label{sub:Multi-parent conflicts}

TODO: alice and bob

TODO: edge and vertex numbers

@@

TODO: plus(x, \hole) that both Alice and Bob move the "x" to the right -> This results in two edges from plus to x.

@@

\subsubsection{Cycles}
\label{sub:Cycles}

@@

Although this case is likely to be infrequent.

Finally, note that through we present several complex scenarios, this is merely for presentation.
In practice, these complex scenarios occur less frequently than portrayed here.

@@

\section{Formalism}
\label{sec:Formalism}

@@

\subsection{Syntax}
\label{sub:Syntax}

@@

\subsection{Convergent Graphs}
\label{sub:Convergent Graphs}

\subsubsection{Graphs}
\label{sub:Graphs}

@@

\subsubsection{Graph Actions}
\label{sub:Graph Actions}

A graph action is a +E or -E.

@@

\subsection{Agda Mechanization}
\label{sub:Agda Mechanization}

@@

\subsection{Edit Actions}
\label{sub:Edit Actions}

Mapping to graph actions

\section{Type System}
\label{sec:Type System}

@@

@@

\section{Implementation}
\label{sec:Implementation}
GRV -- how it is implemented, how it connects to the formalism, describe the graph,

@@

\section{Related Work}
\label{sec:Related Work}

@@

\section{Discussion and Conclusion}
\label{sec:Discussion and Conclusion}

@@

\subsection{Variable names, strings, and numbers}
\label{sub:Variable names, strings, and numbers}

@@

%% Appendix
\appendix
\section{Appendix}
\label{apx:Appendix}
